---
title: "Chapter 2 Data Manipulation and Simple Calculations"
output: html_notebook
---

### First, load the library

```{r}
library(GCDkit)
```

### Exercise 2.1: Subsetting a numeric vector, vector operations

GCDkit contains several built-in datasets, normally needed just for correct system functioning. One of these is atomic weights, stored in the named vector mw. We will use it to show some simple vector operations.

a) Display the whole vector. 
```{r}
mw
```
What is the atomic weight of Rb?
```{r}
mw['Rb']

```
b) What is the average value of the whole vector?
```{r}
mean(mw)
```
c) Which atoms have the atomic weight higher than 170?
```{r}
names(mw)[mw>170]
```
d) Display the names of six lightest and six heaviest elements in the dataset.
```{r}
sort(mw)[1:6]
```
```{r}
rev(sort(mw))[1:6]
```

---

### Excercise 2.2: Loading files, matrix/data frame manipulations

The file sazava.data contains selected major- and trace-element analyses from the ~354 Ma old Sázava suite of the Central Bohemian Plutonic Complex (CBPC; Bohemian Massif, Czech Republic) (Janoušek et al. 2000, 2004).

a) Read analyses stored in the tab-delimited data file into a data frame WR. 

b) Find out the names of available variables (= column names).
```{r}
sazava <- read.table("../Janousek_et_al_2015_modelling_Springer/Part_1/Code/Data/sazava.data", sep="\t")
print(colnames(sazava))
```
c) What is the MgO content of sample Sa-1?
```{r}
sazava["Sa-1","MgO"]
```
d) Show all available numeric data for samples Po-1 and Po-4.
```{r}
sazava <- sazava[,-(1:6)]
sazava[c("Po-1","Po-4"),]
```

e) Calculate the total of the column “Na2O”.

f) Display names of three samples with the lowest and the highest SiO2 contents.

g) Calculate averages of all variables.

h) Display a table with three columns: SiO2, MgO and Na2O/K2O.

